Weak interspecific interactions in a sagebrush steppe? Conflicting evidence from observations and experiments

### 2018-03-23 removals/manuscript/removal_ms.tex
Minor changes to removal ms. final version

@@

All data and R code necessary to reproduce our analysis will be deposited in the Dryad Digital Repository once the manuscript is accepted. The current version of the computer code and all necessary data are available at \texttt{https://github.com/pbadler/ExperimentTests/tree/master/removals}.

@@

In spring 2011, in a large exclosure containing six of the historical permanent quadrats, we established an additional 16 quadrats for the removal treatments. New quadrats were selected to be similar to historical plots, but also needed to have enough of the focal species present so that removal experiments could be conducted. We evaluated potential locations for the new quadrats near the original quadrats and then used only those locations with at least 5\% canopy cover of \textit{Artemisia tripartita} and at least 3\% combined basal cover of the three focal perennial grasses. We rejected locations falling on hill slopes, areas with greater than 20\% bare rock, or with over 10\% cover of the woody shrubs \textit{Purshia tridentata} or \textit{Amelanchier utahensis}. We mapped the new quadrats in June, 2011 and then implemented the removal treatments in September, 2011. We randomly assigned 8 of the sixteen new quadrats to the sagebrush removal treatment, and the remaining 8 to the grass removal treatment. We cut sagebrush stems at ground level and applied a 5\% solution of Roundup Pro (Glyphosate, Monsanto) to the cut stems. In the grass removal plots, we painted the same herbicide solution on all perennial grasses. We removed plants 50 cm beyond the quadrat boundaries to minimize edge effects. The sagebrush removal was virtually 100\% effective; grass removals required additional herbicide applications in April and May of each year, with fewer grasses remaining each year. Appendix S1 Fig. \ref{fig:photos} shows photographs of example quadrats before and after treatment. We repeated mapped censuses of all plots each June from 2012-2016.

@@

Our regression models for survival, growth and recruitment follow previous work \citep{adler_coexistence_2010,chu_large_2015}. The current analysis departs from previous work with the inclusion of a categorical removal treatment effect. Although the three vital rate models differ in functional form (see Appendix A for complete model descriptions), all of them have a linear predictor, $\mu$, containing the following terms:
\begin{equation}
\mu_{ij} = \gamma_j+ \varphi_{jg}+  \chi_j  + \sum \limits_{m} \omega_{jm} {W}_{ijm}
\label{eqn:survReg}
\end{equation}
where $i$ refers to an individual genet (for survival and growth) or quadrat (for recruitment) of species $j$, $\gamma$ is a time-dependent intercept (year effect), $\varphi$ is the coefficient for the effect of spatial location (`$g$' represents a group of nearby quadrats located within one pasture or grazing exclosure), $\chi$ is the removal treatment effect, 
$W_{ijm,t}$ is the crowding on species $j$ exerted by neighbors of species $m$ at the individual (survival and growth) or
quadrat (recruitment) scale, and $\omega_{jm}$ is an interaction coefficient which determines the impact of crowding by species $m$ on the 
focal species $j$. For plants in removal plots in 2012, we set crowding for removed neighbor species to zero, rather than using the pre-treatment (2011) data.

@@

Although all the parameters in our model are estimated with uncertainty, we chose not to propage parameter uncertainty through these 
simulations because we wanted to maximize our ability to detect the population-level consequences of a small removal effect. 
We previously found that interspecific competition is weak relative to the effects of interannual environmental variability 
\citep{Tredennick2017}. Including parameter uncertainty along with random year effects would conflate parameter uncertainty 
with environmental variability, overstating the magnitude of natural fluctuations and making it more difficult to show that competitive release
is larger than the observationally-based model predicts. Therefore, we used estimated means of all parameters, including those with 
confidence intervals that overlapped zero. Based on the same rationale, we used a deterministic version of the model (e.g., 
recruitment is the $\lambda$ of Eqn. \ref{eqn:recrProcessModel}, rather than a Poisson random variable with mean $\lambda$). 
Specifically, we used the recruitment regression to calculate the expected number of new recruits at time $t+1$, which we then multiplied 
by the observed mean size of recruits for the focal species to get expected cover of new recruits.

@@

Our population model based on historical, observational data revealed weak interspecific interactions \citep{adler_coexistence_2010,chu_large_2015}. As a result, the model predicts very little competitive release among the four dominant species in a sagebrush steppe. 
Because the model was based solely on observational
data, we were suspicious of its ability to accurately describe the outcome of interspecific interactions, so we conducted a removal experiment to test it. 
Our approach represents a strong test of a surprising prediction for two reasons. First, the prediction was already published; the experiment 
gave us a very real opportunity to be wrong in public. Second, testing an observationally-based model with an experimental manipulation, 
not just additional observations, further increases the potential to confront the model with a novel situation outside the historical range of variation.
We expected this study to provide clear evidence for the success or failure of our previous model's predictions. Unfortunately, our results do not provide that clarity. In the sections below, we first summarize the evidence that our experiment revealed greater competitive release than our baseline model predicted, then explain weaknesses in the experimental results that complicate the interpretation, and end by discussing general lessons for the study of interspecific interactions.

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% APPENDICES !
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\clearpage 
\newpage

\setcounter{page}{1}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{section}{0}
\setcounter{table}{0}

\centerline{\Large \textbf{Appendix S1}}
\centerline{Adler et al., ``Weak interspecific interactions in a sagebrush steppe? Conflicting evidence from observations and experiments,'' \textit{Ecology}}

\vspace{0.4in}

\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{\arabic{section}}

\section{Detailed Methods} 
\label{AppendixS1}

\subsection{Statistical models of vital rates}

We modeled the survival probability, $S$, of genet $i$ in species $j$ (which is in quadrat group $g$ and removal treatment $h$), 
from time $t$ to $t+1$  as
\begin{equation}
\mbox{logit}(S_{ij,t+1}) = \gamma_{j,t}^S + \varphi_{jg}^S+  \chi_{jh}^S  + \beta_{j,t}^S u_{ij,t} +  
\sum \limits_{m} \omega_{jm}^S {W}_{ijm,t}
\label{eqn:survReg}
\end{equation}
where $\gamma$ is a time-dependent intercept (year effect), $\varphi$ is the coefficient for the 
effect of spatial location group, $\chi$ is the removal treatment effect,and  $\beta$ is the coefficient that 
represents the effect of log genet size, $u$, on survival. 
$W_{ijm,t}$ is the crowding of individual $i$ in species $j$ by neighbors of species $m$ (explained below), and 
$\omega_{jm}$ is an interaction coefficient which determines the impact of crowding by species $m$ on the focal species $j$.

Our growth model has a similar structure. The change in genet size from time $t$ to $t+1$ , conditional on survival, is given by:
\begin{equation}
u_{ij,t+1} = \gamma_{j,t}^G + \varphi_{jg}^G+  \chi_{jh}^G  + \beta_{j,t}^G u_{ij,t} + 
\sum \limits_{m} \omega_{jm}^G {W}_{ijm,t} + \varepsilon_{ij,t}^G .
\label{eqn:growReg}
\end{equation}
To capture the non-constant error variance in growth, we modeled the variance $\varepsilon$ about the growth 
curve (\ref{eqn:growReg}) as a nonlinear function of predicted genet size:
\begin{equation}
Var(\varepsilon_{ij,t}^G) = a \exp(b\hat{u}_{ij,t+1}).
\label{eqn:growVar}
\end{equation}
where $\hat{u}_{ij,t+1}$ is the right-hand side of (\ref{eqn:growReg}) without $\varepsilon$.

Our model includes crowding from the four dominant species, \textit{A. tripartita}, \textit{H. comata}, \textit{P. spicata}, and \textit{P. secunda}, as well as two covariates representing total crowding from (1) all other perennial grasses and shrubs, which were mapped as polygons, and (2) all forb species, which were mapped as points. The next section (\textit{Modeling local crowding}) provides a detailed explanation of how we modeled crowding based on the size of and distance to neighboring plants around each focal genet.

We model recruitment at the quadrat level rather than at the individual genet level because the mapped data do not allow 
us to determine which recruits were produced by which potential parent genets. We assume that the number of individuals, $y$, of species $j$ recruiting at 
time $t+1$ in the location $q$ follows a negative binomial distribution:
\begin{equation}
y_{jq,t+1}= NegBin(\lambda_{jq,t+1},\theta) 	   
\label{eqn:recrDataModel}
\end{equation}
where $\lambda$ is the mean and $\theta$ is the size parameter (\texttt{dnegbin(p,$\theta$)} in the BUGS language, with $p=\theta/(\theta+\lambda)$). 
In turn, $\lambda$ depends on the composition of the quadrat in the previous year:
\begin{equation}
\lambda_{jq,t+1} = C'_{jq,t} \exp{\left(\gamma_{j,t}^R +  \varphi_{jg}^R + \chi_{jh}^R + 
\sum \limits_{m} \omega_{jm}^R \sqrt{C'_{mq,t}} \right)}
\label{eqn:recrProcessModel}
\end{equation}
where the superscript $R$ refers to Recruitment, $C'_{jq,t}$ is the `effective cover' (cm$^2$) of species $j$ in quadrat $q$ at time $t$, $\gamma$ 
is a time-dependent intercept, $\varphi$ is a coefficient for the effect of spatial location group,
$\chi$ is the removal treatment effect, and $\omega_{jm}$ is the coefficient that determines the effect of species $m$ cover on recruitment by species $j$. 
Eqn (\ref{eqn:recrProcessModel}) is essentially the Ricker model for discrete time population growth \citep{ricker_stock_1954}, except that we use the square root of cover in the density-dependent term. We previously used a Ricker equation \citep{adler_coexistence_2010}, but using the square-root transformation instead gave a lower Deviance Information Criterion (DIC) value \citep{spiegelhalter_bayesian_2002} for the recruitment model. Following previous work \citep{adler_coexistence_2010}, we treated year and spatial group variables as random factors, allowing intercepts to vary among years and spatial locations.

Due to the possibility that plants outside the mapped quadrat could contribute recruits to the focal quadrat or interact with plants in the focal quadrat, we estimated effective cover as a mixture of the observed cover, $C$, in the focal quadrat, $q$, and the mean cover, $\bar{C}$, across the spatial location group, $g$, in which the 
quadrat is located: $C'_{jq,t}=p_j C_{jq,t}+(1-p_j) \bar{C}_{jg,t}$, where $p$ is a mixing fraction between 0 and 1 that was estimated as part of fitting the model.

We ran Markov Chain Monte Carlo (MCMC) simulations in WinBUGS 1.4 \citep{lunn_winbugs_2000} to estimate the recruitment model parameters. Each model was run for 20,000 iterations and two independent chains with different initial values for parameters. We discarded the initial 10,000 samples. Convergence was observed graphically for all parameters, and confirmed by the Brooks-Gelman statistic \citep{brooks_general_1998}. For all three vital rate models, we considered the effect of removal treatments, or other coefficients of interest, to be significant when the 95\% credible interval (or posterior interval) on the estimate of $\chi$ did not overlap zero.

To incorporate the fitted recruitment function into the IPM, we assumed that individual fecundity increases 
linearly with individual area (equal to $e^u$ for a size-$u$ individual). The recruitment kernel is thus
$R_j(v,u,\vec{n})=c_{0,j}(v)e^{u}\Phi_j$ \citep{adler_coexistence_2010} where $c_{0,j}$ is the
size distribution of new recruits. $\Phi_j$ is calculated from eqn. (\ref{eqn:recrDataModel}) with total cover values computed from the population densities as $C'_{j,t} = \int e^u n_j(u,t) du$.

\subsection{Modeling local crowding}
We modeled the crowding experienced by a focal genet in each year $t$ as a function of the distance to and size of neighbor genets. In previous work, 
we assumed that the decay of crowding with neighbor distance followed a Gaussian function \citep{chu_large_2015}, but here we use a data-driven 
approach \citep{teller_linking_2016}. We model the crowding experienced by genet $i$ of species $j$ from neighbors of species $m$ as the sum of neighbor 
areas across a set of concentric annuli, $k$, centered at the plant,
\begin{equation}
w_{ijmk,t} = F_{jm}(d_{k})A_{ijmk,t}     
\label{eqn:wik}
\end{equation}
where $F_{jm}$ is the competition kernel (described below) for effects of species $m$ on species $j$, 
$d_{k}$ is the average of the inner and outer radii of annulus $k$, 
and $A_{ijmk,t}$ is the total area of genets of species $m$ in annulus $k$ around the focal genet in year $t$. The total crowding on 
the focal genet exerted by species $m$ is
\begin{equation}
W_{ijm,t}  =\sum_k {w_{ijmk,t}} .
\label{eqn:wijm}
\end{equation} 
Note that $W_{ijj,t}$ gives intraspecific crowding. The $W$'s are then 
included as covariates in the survival (\ref{eqn:survReg}) and growth (\ref{eqn:growReg}) regressions.

We assume that competition kernels $F_{jm}(d)$ are non-negative and decreasing, so that distant plants have less effect 
than close plants. Otherwise, we let the data dictate the shape of the kernel by fitting a spline model 
using the methods of \citet{teller_linking_2016}. The shape of $F_{jm}$ is determined by a set of spline basis coefficients $\vec{b}_{jm}$
and a smoothing parameter $\eta$ that controls the complexity of the fitted kernel. 
Demographic models such as \eqref{eqn:survReg} then have $\gamma$, $\varphi$, $\chi$ , 
$\beta$, $\boldsymbol{\omega}$, $\vec{b}$ and $\eta$ as parameters to be fitted. We implemented this in the statistical computing environment, \texttt{R}, 
by making the spline coefficients and $\eta$ the arguments of an objective function that calculates  each $w_{ijmk,t}$ value using the input spline coefficients, 
calls the model-fitting functions \texttt{lmer} and/or \texttt{glmer} to fit the other parameters in the survival and growth regressions, 
and returns an approximate AIC value and model degrees of freedom ($df$) for survival and growth combined. We used the $\vec{b}$ values at the smoothest 
(smallest $df$) local minimum of AIC as a function of $df$, as in \cite{teller_linking_2016}. This approach assumes that one measure of crowding affects 
survival and growth. In addition, for fitting the kernels we assumed that survival and growth depended only on intraspecific crowding, and thus only fitted the
within-species competition kernels $F_{jj}$. Based on previous work with parametric kernels, \citep{adler_coexistence_2010}, we know that interspecific effects are rarely significantly different from zero, so there is not enough signal to fit interspecific kernels nonparametrically. Therefore, we set all $F_{mj}$ equal to $F_{jj}$, meaning that 
the within-species competition kernel for species $j$ is also used to determine the distance-dependence of species $j$ neighbors on all other species. 
We used data from all historical plots and contemporary control-treatment plots to estimate the competition kernels, which are shown in Fig. \ref{fig:CompKernels}.

Once we had estimated the competitions kernels, we used them to calculate the $w_{ijmk,t}$ values for each individual in each year, 
and fitted the full survival and growth regressions, which include the interspecific interaction coefficients, $\omega_{jm}$. 
All genets in a quadrat were included in calculating the $w$'s, but plants located within 5 cm of quadrat edges were not used in fitting the regressions. 
We fit the models using the \texttt{R-INLA} package for R \citep{rue_approximate_2009}. We used the same approach when exploring variations 
on the growth regressions, such as the addition of a year-by-treatment interaction or information about the locations of 
individual grasses relative to removed \textit{A. tripartita}. We also compared the species composition of the control and removal plots at both 
the quadrat and neighborhood ($W$) scales.

\subsection{Interspecific covariance in local crowding} 
We explored interspecific covariance in local crowding experienced by individual plants, by regressing the $W$ values exerted by one neighbor species,
 the response variable, against the $W$ values of all other species, the independent variables. Because some $W=0$, we conducted two 
 separate regressions. First, using all $W$'s, we fitted a generalized linear model with a logit link function to evaluate whether the 
 probability that the focal species' $W$ equals 0 is influenced by the value of other species' $W$'s. In this model, the dependent 
 variable is a Bernoulli variate coding for zero versus non-zero value of the focal species' crowding, and the independent 
 variables are the $W$'s for all other species. Second, for the set of records in which the focal species has $W>0$, we 
 performed a linear regression, where the focal species' $W$ is the dependent variable, and the other species' $W$'s are 
 the independent variables. We repeated these regressions for each focal species. Due to large samples size, interspecific $W$ 
 values were often statistically significant predictors of intraspecific. However, they explained very little variance. The 
 maximum reduction in deviance for the generalized linear regressions and $R^2$ for the linear regressions were both less 
 than 8\%. The \texttt{R} code for this analysis is included as ..\texttt{\textbackslash Wdistrib\textbackslash exploreSurvivalWs.r}.

\subsection{Mean field approximation of local crowding for the IPM} 
\label{sec:kernelMethods} 
\citet{adler_coexistence_2010} developed a mean field approximation for local crowding when the
competition kernels are all Gaussian functions, $F_{jm}(d) = e^{-\alpha_{jm} d^2}$. The approximation is explained in 
the online SI to \citet{adler_coexistence_2010} and in section 5.3 of \citet{Ellner2016}. 
Here we explain how that approximation was modified for the IPMs in this paper, which
used fitted nonparametric competition kernels.

The mean field approximation is based on the observed spatial distribution patterns of the species \citep{adler_coexistence_2010}. 
In both the observed data and IBM simulations, heterospecific individuals were approximately randomly distributed with respect to each other, 
but conspecific individuals displayed a non-random, size-dependent patterns: small genets were randomly distributed, while large genets 
were segregated from each other. The overdispersion of large conspecific genets is incorporated into the IPM 
with a `no-overlap' rule explained below, as in \citep{adler_coexistence_2010}.

For $j \ne m$ (between-species competition), overlap between individuals is allowed. The mean field approximation is 
that from the perspective of any focal plant in species $j$, individuals of species $m$ are distributed at random in space, 
independent of each other and of their size.

Consider the region between the circles of radius $x$ and $x+dx$ centered on a focal genet of species $j$. The area of this annulus
is $2 \pi x \; dx$  to leading order for $dx \approx 0$. A species $m$ genet 
in the annulus puts competitive pressure $F_{jm}(x)$ times its area on
the focal genet. The expected total competitive pressure from all such genets 
is therefore is $F_{jm}(x) 2 \pi x \; dx$ times the expected fractional cover of species $m$ in the annulus 
(fractional cover is the total area of species $m$ genets, as a fraction of the total area). The excepted fractional cover $C_m$ of species $m$
in the annulus equals its fractional cover in the habitat as a whole, because of the assumption of random distribution
spatial distributions. We therefore have $C_m  = \int e^u n_m(u,t) du/A$ where $A$ is the total area of the habitat. 
The total expected competitive pressure on a species-$j$ genet due to species $m$ is then 
\begin{equation}
W_{jm} = \int_0^\infty{C_m F_{jm}(x) 2 \pi x \; dx}  = C_m \left [2 \pi \int_0^{\infty} x F(x) \, dx \right ].
\label{eqn:wbarm}
\end{equation} 
The quantity in square brackets is a constant (that is, it only depends on what the kernel function
is) so it can be computed once and for all for each kernel used in the IPM. The integral is finite because
all fitted kernels fall to zero at a finite distance from the focal plant.

Our kernel fitting method only uses competition kernel values at the ``mid-ring'' distances
halfway between the inner and outer radii of a series of annuli around each focal
plant, scaled so that the value at the innermost mid-ring distance equals 1. 
In the IPM we defined the kernel at other distances by linear interpolation between values at 
mid-ring distances, except that for the innermost ring a kernel value of 1 was specified at the
outer radius of the ring and at distance $x=0$.

Now consider within-species competition. We assume that conspecifics cannot overlap. Genet shapes are irregular, but we 
nonetheless implement the no-overlap rule by assuming that a genet of log area $u_i$ is a 
circle of radius $r_i$ where $\pi r_i^2 = e^{u_i}$. The no-overlap rule is then that the centroids of two conspecific individuals 
must be separated by at least the sum of their radii.

For any one focal genet, the no-overlap restriction on its neighbors' locations affects 
only a negligibly small part of the habitat. The expected cover of individuals in the places
where they can occur (relative to one focal plant) is thus assumed to equal their expected locations
in the habitat as a whole.

Let $C_m(u)$ be the total cover of species $m$ genets of radius $r$ or smaller, 
\begin{equation}
C_m(r) = \int_L^{\log(\pi r^2)}{\! \! \! e^z n_m(z,t) \, dz} .
\label{eqn:cm}
\end{equation}
A focal genet of radius $r$ cannot have any conspecific neighbors centered 
at distances less than $r$. It can have a neighbor centered at distance $x>r$ if that neighbor's
radius is no more than $x-r$. Adding up the expected cover of all such possible neighbors
for a focal genet of radius $r$,    
\begin{equation}
W_{mm}(r) = 2 \pi \int_r^{\infty}F_{mm}(x) x C_m(x-r) \, dx
\label{eqn:wbarmr} 
\end{equation}
This integral is again finite and computable because the kernels $F$ fall to 0 at finite $x$.

\clearpage 
\newpage

\section{Supplementary Tables}

% supplementary tables
\begin{table}[h]
\caption{Statistical models of year-to-year changes in log(cover) for the four focal species. Values in square brackets show 95\%
 credible intervals.}
\centering
%\begin{center}
\begin{tabular}{l c c c c }
\hline
Species & \textit{A. tripartita} & \textit{H. comata} & \textit{P. secunda} & \textit{P. spicata} \\
\hline
(Intercept)           & $-0.03$          & $0.08$           & $-0.14$          & $0.02$           \\
                      & $[-0.48;\ 0.41]$ & $[-0.14;\ 0.30]$ & $[-0.56;\ 0.27]$ & $[-0.24;\ 0.29]$ \\
TreatmentNo\_grass    & $-0.05$          &                  &                  &                  \\
                      & $[-0.74;\ 0.64]$ &                  &                  &                  \\
TreatmentNo\_shrub    &                  & $0.24$           & $-0.05$          & $0.18$           \\
                      &                  & $[-0.12;\ 0.60]$ & $[-0.35;\ 0.25]$ & $[-0.01;\ 0.38]$ \\
\hline
%AIC                   & 227.74           & 114.65           & 205.41           & 168.17           \\
%BIC                   & 240.01           & 124.69           & 218.18           & 181.40           \\
Log Likelihood        & -108.87          & -52.33           & -97.70           & -79.09           \\
Num. obs.             & 86               & 55               & 95               & 104              \\
Num. groups: quad     & 19               & 11               & 21               & 22               \\
Num. groups: year     & 5                & 5                & 5                & 5                \\
Var: quad (Intercept) & 0.44             & 0.00             & 0.02             & 0.00             \\
Var: year (Intercept) & 0.00             & 0.02             & 0.19             & 0.07             \\
Var: Residual         & 0.52             & 0.36             & 0.38             & 0.24             \\
\hline
\multicolumn{5}{l}{\scriptsize{$^*$ 0 outside the credible interval}}
\end{tabular}
\label{table:coefficients}
%\end{center}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:32:17 2016
\begin{table}[ht]
\centering
\caption{Summary of posterior distributions for fixed effects for the \textit{A. tripartita} survival model. ``logarea" is the effect of plant size, ``Treatment*" is the removal effect, and the ``W.*" coefficients are effects of neighborhood crowding (the $\omega$s in eqn. \ref{eqn:survReg}). ``quant" refers to quantile and ``kld" reports the Kullback-Leibler divergence between the Gaussian and the (simplified) Laplace approximation to the marginal posterior densities. } 
\label{ARTRsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & -0.2613 & 0.1344 & -0.5276 & -0.2605 & 0.0001 & -0.2588 & 0.0000 \\ 
  logarea & 0.7093 & 0.0400 & 0.6315 & 0.7090 & 0.7885 & 0.7085 & 0.0000 \\ 
  TreatmentNo\_grass & -0.9994 & 0.6453 & -2.2156 & -1.0182 & 0.3228 & -1.0565 & 0.0000 \\ 
  W.ARTR & -2.8522 & 0.4187 & -3.6968 & -2.8443 & -2.0513 & -2.8284 & 0.0000 \\ 
  W.HECO & -0.0720 & 0.0615 & -0.1856 & -0.0747 & 0.0565 & -0.0804 & 0.0000 \\ 
  W.POSE & 0.0617 & 0.0895 & -0.1061 & 0.0588 & 0.2459 & 0.0528 & 0.0000 \\ 
  W.PSSP & 0.0387 & 0.0475 & -0.0500 & 0.0370 & 0.1369 & 0.0335 & 0.0000 \\ 
  W.allcov & -0.0065 & 0.0109 & -0.0282 & -0.0065 & 0.0148 & -0.0063 & 0.0000 \\ 
  W.allpts & 0.1293 & 0.1653 & -0.1956 & 0.1294 & 0.4533 & 0.1296 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:33:00 2016
\begin{table}[ht]
\centering
\caption{Summary of posterior distributions for fixed effects for the \textit{H. comata} survival model. See Table \ref{ARTRsurvival} for an explanation of coefficient names
and column headers.} 
\label{HECOsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.4753 & 0.2011 & 1.0764 & 1.4753 & 1.8736 & 1.4753 & 0.0000 \\ 
  logarea & 1.1991 & 0.0747 & 1.0622 & 1.1954 & 1.3561 & 1.1874 & 0.0000 \\ 
  TreatmentNo\_shrub & 0.2915 & 0.3900 & -0.4693 & 0.2898 & 1.0610 & 0.2864 & 0.0000 \\ 
  W.ARTR & -0.0063 & 0.0030 & -0.0121 & -0.0063 & -0.0005 & -0.0063 & 0.0000 \\ 
  W.HECO & -1.6962 & 0.1458 & -1.9894 & -1.6939 & -1.4163 & -1.6891 & 0.0000 \\ 
  W.POSE & 0.0649 & 0.0563 & -0.0442 & 0.0644 & 0.1768 & 0.0634 & 0.0000 \\ 
  W.PSSP & 0.0313 & 0.0360 & -0.0381 & 0.0309 & 0.1031 & 0.0301 & 0.0000 \\ 
  W.allcov & -0.0012 & 0.0062 & -0.0135 & -0.0012 & 0.0110 & -0.0012 & 0.0000 \\ 
  W.allpts & -0.1295 & 0.1004 & -0.3273 & -0.1293 & 0.0670 & -0.1290 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:34:30 2016
\begin{table}[ht]
\centering
\caption{Summary of posterior distributions for fixed effects for the \textit{P. secunda} survival model. See Table \ref{ARTRsurvival} for an explanation of coefficient names
and column headers.} 
\label{POSEsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.4020 & 0.1922 & 1.0176 & 1.4031 & 1.7793 & 1.4050 & 0.0000 \\ 
  logarea & 1.0362 & 0.0645 & 0.9139 & 1.0343 & 1.1696 & 1.0307 & 0.0000 \\ 
  TreatmentNo\_shrub & -0.1604 & 0.1622 & -0.4784 & -0.1606 & 0.1583 & -0.1610 & 0.0000 \\ 
  W.ARTR & 0.0000 & 0.0020 & -0.0038 & 0.0000 & 0.0039 & -0.0000 & 0.0000 \\ 
  W.HECO & -0.0224 & 0.0208 & -0.0631 & -0.0225 & 0.0185 & -0.0227 & 0.0000 \\ 
  W.POSE & -2.0073 & 0.1112 & -2.2290 & -2.0062 & -1.7921 & -2.0038 & 0.0000 \\ 
  W.PSSP & 0.0468 & 0.0238 & 0.0004 & 0.0466 & 0.0939 & 0.0463 & 0.0000 \\ 
  W.allcov & -0.0061 & 0.0035 & -0.0130 & -0.0061 & 0.0008 & -0.0061 & 0.0000 \\ 
  W.allpts & -0.0383 & 0.0567 & -0.1494 & -0.0384 & 0.0730 & -0.0385 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.2.2 by xtable 1.8-2 package
% Mon Sep 19 08:35:48 2016
\begin{table}[ht]
\centering
\caption{Summary of posterior distributions for fixed effects for the \textit{P. spicata} survival model. See Table \ref{ARTRsurvival} for an explanation of coefficient names
and column headers.} 
\label{PSSPsurvival}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 1.1978 & 0.1639 & 0.8783 & 1.1962 & 1.5257 & 1.1933 & 0.0000 \\ 
  logarea & 1.5138 & 0.0935 & 1.3336 & 1.5122 & 1.7029 & 1.5092 & 0.0000 \\ 
  TreatmentNo\_shrub & -0.2466 & 0.2023 & -0.6454 & -0.2460 & 0.1489 & -0.2449 & 0.0000 \\ 
  W.ARTR & 0.0099 & 0.0022 & 0.0057 & 0.0099 & 0.0142 & 0.0099 & 0.0000 \\ 
  W.HECO & 0.0025 & 0.0289 & -0.0550 & 0.0028 & 0.0586 & 0.0032 & 0.0000 \\ 
  W.POSE & 0.0517 & 0.0389 & -0.0222 & 0.0508 & 0.1308 & 0.0490 & 0.0000 \\ 
  W.PSSP & -1.1267 & 0.0739 & -1.2743 & -1.1258 & -0.9838 & -1.1242 & 0.0000 \\ 
  W.allcov & 0.0131 & 0.0037 & 0.0060 & 0.0131 & 0.0203 & 0.0131 & 0.0000 \\ 
  W.allpts & 0.0756 & 0.0658 & -0.0540 & 0.0757 & 0.2045 & 0.0760 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

% cross reference to appendix
\usepackage{xr}
\externaldocument{appendix}

@@

All data and R code necessary to reproduce our analysis are available on the Dryad Digital Repository (insert url here).

@@

In spring 2011, in a large exclosure containing six of the historical permanent quadrats, we established an additional 16 quadrats for the removal treatments. New quadrats were selected to be similar to historical plots, but also needed to have enough of the focal species present so that removal experiments could be conducted. We evaluated potential locations for the new quadrats near the original quadrats and then used only those locations with at least 5\% canopy cover of \textit{Artemisia tripartita} and at least 3\% combined basal cover of the three focal perennial grasses. We rejected locations falling on hill slopes, areas with greater than 20\% bare rock, or with over 10\% cover of the woody shrubs \textit{Purshia tridentata} or \textit{Amelanchier utahensis}. We mapped the new quadrats in June, 2011 and then implemented the removal treatments in September, 2011. We randomly assigned eight of the 16 new quadrats to the sagebrush removal treatment, and the remaining eight to the grass removal treatment. We cut sagebrush stems at ground level and applied a 5\% solution of Roundup Pro (Glyphosate, Monsanto) to the cut stems. In the grass removal plots, we painted the same herbicide solution on all perennial grasses. We removed plants 50 cm beyond the quadrat boundaries to minimize edge effects. The sagebrush removal was virtually 100\% effective; grass removals required additional herbicide applications in April and May of each year, with fewer grasses remaining each year. Appendix S1: Fig. \ref{fig:photos} shows photographs of example quadrats before and after treatment. We repeated mapped censuses of all plots each June from 2012-2016.

@@

Our regression models for survival, growth and recruitment follow previous work \citep{adler_coexistence_2010,chu_large_2015}. The current analysis departs from previous work with the inclusion of a categorical removal treatment effect. Although the three vital rate models differ in functional form (see Appendix S1 for complete model descriptions), all of them have a linear predictor, $\mu$, containing the following terms:
\begin{equation}
\mu_{ij} = \gamma_j+ \varphi_{jg}+  \chi_j  + \sum \limits_{m} \omega_{jm} {W}_{ijm}
\label{eqn:survReg}
\end{equation}
where $i$ refers to an individual genet (for survival and growth) or quadrat (for recruitment) of species $j$, $\gamma$ is a time-dependent intercept (year effect), $\varphi$ is the coefficient for the effect of spatial location (`$g$' represents a group of nearby quadrats located within one pasture or grazing exclosure), $\chi$ is the removal treatment effect, 
$W_{ijm,t}$ is the crowding on species $j$ exerted by neighbors of species $m$ at the individual (survival and growth) or
quadrat (recruitment) scale, and $\omega_{jm}$ is an interaction coefficient which determines the impact of crowding by species $m$ on the 
focal species $j$. For plants in removal plots in 2012, we set crowding for removed neighbor species to zero, rather than using the pre-treatment (2011) data.

@@

Although all the parameters in our model are estimated with uncertainty, we chose not to propage parameter uncertainty through these 
simulations because we wanted to maximize our ability to detect the population-level consequences of a small removal effect. 
We previously found that interspecific competition is weak relative to the effects of interannual environmental variability 
\citep{Tredennick2017}. Including parameter uncertainty along with random year effects would conflate parameter uncertainty 
with environmental variability, overstating the magnitude of natural fluctuations and making it more difficult to show that competitive release
is larger than the observationally-based model predicts. Therefore, we used estimated means of all parameters, including those with 
confidence intervals that overlapped zero. Based on the same rationale, we used a deterministic version of the model (e.g., 
recruitment is the $\lambda$ of Appendix S1: Eqn. \ref{eqn:recrProcessModel}, rather than a Poisson random variable with mean $\lambda$). 
Specifically, we used the recruitment regression to calculate the expected number of new recruits at time $t+1$, which we then multiplied 
by the observed mean size of recruits for the focal species to get expected cover of new recruits.

@@

Our population model based on historical, observational data revealed weak interspecific interactions \citep{adler_coexistence_2010,chu_large_2015}. As a result, the model predicts very little competitive release among the four dominant species in a sagebrush steppe. 
Because the model was based solely on observational
data, we were suspicious of its ability to accurately describe the outcome of interspecific interactions, so we conducted a removal experiment to test it. 
Our approach represents a strong test of a surprising prediction for two reasons. First, the prediction was already published, so the experiment 
gave us a very real opportunity to be wrong in public. Second, testing an observationally-based model with an experimental manipulation, 
not just additional observations, further increases the potential to confront the model with a novel situation outside the historical range of variation.
We expected this study to provide clear evidence for the success or failure of our previous model's predictions. Unfortunately, our results do not provide that clarity. In the sections below, we first summarize the evidence that our experiment revealed greater competitive release than our baseline model predicted, then explain weaknesses in the experimental results that complicate the interpretation, and end by discussing general lessons for the study of interspecific interactions.

@@
